VOI-aware Monte Carlo Sampling in Trees

### 2011-12-08 paper/vago/paper.tex
experiments on bandits

One way to account for this re-use would be to incroprorate the
`future' value of information into a VOI estimate. However, this 
approach appears to be complicated. Alternatively, one can behave
myopically in search tree depth:
\begin{enumerate}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-termination-condition});
\item save the unused sample budget for search in future state, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposedly a
fixed value for a given problem type and algorithm implementation would
work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
confirms that a learned fixed cost and sample redistribution
substantially improve the performance of the VOI-based sampling policy
in game tree search.

@@

Figure~\ref{fig:random-instances}

@@

>>>>

One way to account for this re-use would be to incroprorate the
`future' value of information into a VOI estimate. However, this 
approach appears to be complicated. Alternatively, one can behave
myopically in search tree depth:
\begin{enumerate}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-termination-condition});
\item save the unused sample budget for search in future state, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposedly
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

@@

@@

The sampling policies are first compared on random multi-armed bandit 
problem instances, where the Monte Carlo Tree Search algorithm
is reduced to the sample selection policy at the first step: UCT is 
actually UCB1. In the problem of pure exploration in multi-armed bandits
\cite{Bubeck.pure}, the smaller is the exploration factor, the higher
is the expected simple regret; indeed, UCT exhibits the lowest simple
regret in the experiments for the default exploration factor $Cp=\frac
{\sqrt 2} 2$ \cite{Kocsis.uct}.

Figure~\ref{fig:random-instances} shows experiment results for
randomly-generated multi-armed bandits with 32 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range $[0,
  1]$, for a range of sample budgets $32..2048$, with multiplicative
step of $2$. The experiment for each number of samples was repeated
20000 times. UCT is worse than any of the VOI-aware sampling policies,
for example, for 384 samples the simple regret of UCT is $\approx 40$
times greater than of VCT, and $\approx 100$ times greater than of
ECT. On average, it takes $2.5$ times as many samples for UCT to reach
the regret of VCT.
